\mytitle{}

### 2019-07-29 paper/infocom.tex
Somehow reduced to the Infocom page limit...

%?? Discuss the most important conclusions before the outline.
We find that the existing RL work for this task fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
To remedy this, our agents make throttling decisions per-source, and show how an action model built on domain-specific knowledge can increase performance.
This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring (but minimising) the use of software-defined networks (SDN).
%In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
This improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that the use of domain knowledge achieves excellent performance at high host density.
Crucially, the models are protocol- and content-agnostic to offer future-proofing against the rollout of future protocols.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
%	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item A comparison of the effectiveness of different RL algorithms in these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
%	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this by exploiting operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through volume of requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent today.
\emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while leaving each participant anonymous.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning concerned with training an agent to choose an optimal sequence of actions for a given task \cite{RL2E}.
We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is best described as a \emph{Markov Decision Process} (MDP), where agents aim to maximise the \emph{expected reward} received.
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
The policy is refined as more decisions and rewards become available, learning online.
In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new training corpus.

@@

Computing the approximate value of every action available in the current state forms the basis of a policy.
To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy).

@@

\subsection{Motivation}\label{sec:motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
Misclassification can be a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhap allow forgiveness of mistakenly punished flows.
This relies upon a flow-by-flow view of the state space, but by combining knowledge of current state with the last applied action, a punished flow which falls off identically to a legitimate flow can be rescued.

%Which features might be best suited to this problem?
%?? Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
Aggregate network load observed at various locations suggests the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks in progress \cite{DBLP:conf/dsn/LeeKSPY17}.
Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes  which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.
%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)
%?? If we assume amplification attacks, we know it won't be `random' source IPs (since it's mostly-legit servers who think that they're doing a good job by replying)

%\section{A Plan, of Sorts}
%
%\begin{enumerate}
%	\item The main case for contribution in what I have so far:
%	\begin{itemize}
%		\item Past work reliant on unrealistic network models: tcp-like behaviour (and its effects on collateral damage) not captured, disjoint ranges of traffic distribution (no benign heavy-hitters), ISP-like topology.
%		\item I offer more realistic network emulation environment, better treatment of protocol/traffic characteristics.
%	\end{itemize}
%	\item Forthcoming: rethinking state/action spaces to operate at a finer level of granularity. New network model (live tcp back-and-forth), allows us to test collateral damage assumptions in a more realistic manner (and show clear case for moving beyond work of malialis and kudenko)
%\end{enumerate}

%\subsection{Threat Model}
%An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
%Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
%We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
%The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
%The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.
%
%We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
%While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
%The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
%These are APT-level capabilities, whose exploration presents a rich source for future work.

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
%Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
%However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

\subsection{System Design and Assumptions}
Agents are deployed in a network, with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, $U_s$.

@@

\subsubsection{Per-tile updates}
To adjust the value of each activated tile with a different magnitude and direction, we found it effective to compute a different temporal difference value \emph{for each tiling}.
This corrects tilings which contribute too much or too little value without disturbing estimates which are already close to correct.
While we make use of the sum of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to set $\alpha$ to a higher value without divergence.
Compared to the standard algorithm, convergence on effective estimates occurs sooner.

@@

\begin{figure}
	\centering
	\resizebox{0.65\linewidth}{!}{
		\begin{tikzpicture}[
		texts/.style = {text=black},
		labeltexts/.style = {text=gray},
		treeline/.style = {draw=uofgburgundy},
		treenode/.style = {texts, circle, centered, fill=white, treeline},
		load/.style = {fill=uofgcobalt},
		loadline/.style = {draw=uofgcobalt, line width=0.75mm},
		loadhide/.style = {fill=uofgcobalt!40!white},
		external/.style = {fill=uofgrust},
		externalhide/.style = {fill=uofgrust!40!white},
		hideline/.style = {draw=uofgsandstone!40!white},
		hidenode/.style = {treenode, hideline},
		grow'=right
		]
		\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
		\node [treenode, loadhide, right = of mainagent] (inner0) {};
		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner0] (inner1) {};
		\node [hidenode, below right = 0.5cm and 1cm of inner0] (inner2) {};
		\node [treenode, loadhide, below right = 0.5cm and 1cm of inner1] (inner3) {};
		\node [treenode, loadhide, right = of inner3] (inner4) {};
		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
		\node [hidenode, below right = 0.5cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};

		\node [hidenode, above left = 0.5cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
		\node [hidenode, below left = 0.5cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};

@@

Our state space combines global state (network link loads) with per-flow measurements.
Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.

%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}): the bandwidth measurements regularly received from monitors in the environment.
For any flow, an agent then computes the path it would take through the network.
The incoming load along the first hop, last hop, and tertiles of the path may then be tile-coded together.
In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
\Cref{fig:global-state-path} illustrates the process.

Global state built in this way is compatible with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers up to transit ASes.
Computing the path from agent to destination is not computationally expensive.
Multipath routing is fast since typical ECMP routing algorithms simply hash a packet's flow tuple and are deterministic to provide consistent quality-of-service.

@@

We describe and analyse the per-flow features in the state vector throughout \cref{sec:rethinking-the-state-space}.
Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.

@@

While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
End-host servers may make use of canary flows or other active measurement schemes, or employ existing quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
ASes and transit networks may make use of reports received from downstream networks, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
We leave the use of such QoE metrics as future work.
Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training to transfer to reality.

@@

\subsection{Action Space}
When monitoring a flow, an agent uses its state vector to decide which portion of its \emph{inbound} traffic should be dropped.
%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
Furthermore, we need not consider burstiness, fairness or implementation costs (such as bucket sizes and tuning) which could limit scalability.
We offer two models, which differ only on how to choose $p$.
The intuition behind these designs is that an agent may be able to directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may be able to reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).

\subsubsection{Instant control}
Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set.
These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).

\subsubsection{Guarded control}
\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations by constraining how an agent treats each flow via simple finite state automata: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.

We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing.

To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), set this low due to TCP's higher prevalence.
An agent must now choose to punish a flow multiple times in succession to remove its effects, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.

>>>>

%?? Discuss the most important conclusions before the outline.
We find that the existing RL work for this task fails to account for congestion-aware traffic (i.e., TCP, QUIC) and environments with high host density at ingress points.
To remedy this, our agents make throttling decisions per-source, and show how an action model built on domain-specific knowledge can increase performance.
This requires additional engineering: updating RL agents from multiple traces per timestep, smart work scheduling, and an architecture featuring (but minimising) the use of software-defined networks (SDN).
%In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
This improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that the use of domain knowledge achieves excellent performance at high host density.
Crucially, the models are protocol- and content-agnostic to offer future-proofing against the rollout of future protocols.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity models for RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models) and suitable features (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item A description of the system architecture which supports the real-world deployment of such agents (\cref{sec:system-architecture}).
%	\item An investigation of suitable features for this task (\cref{sec:rethinking-the-state-space}).
%	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item A performance comparison of different RL algorithms within these agents, and against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
%	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, to inflict financial harm or as an act of vandalism.
Attackers achieve this by exploiting system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or by overwhelming their target through requests or traffic (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Volume-based attacks are the most relevant to our work, with two varieties being especially prevalent.
\emph{Amplification attacks} exploit services who send large replies in response to small requests, with UDP-based services like DNS and NTP seeing most abuse \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed using the victim's address, causing many large replies to be sent to the intended target---increasing a botnet's throughput while keeping participants anonymous.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning concerned with training an agent to choose an optimal sequence of actions for a given task \cite{RL2E}.
We assume at any point in time $t$, an agent knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is a \emph{Markov Decision Process} (MDP), where agents aim to maximise the \emph{expected reward} received.
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken.
The policy is refined as more decisions and rewards become available, learning online.
In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new training corpus.

@@

Computing the approximate value of every action available in the current state forms the basis of a policy.
To encourage early exploration, actions with maximal value can be chosen each time except when taking random actions with probability $\epsilon$---an \emph{$\epsilon$-greedy policy}).
\Cref{fig:tilecode-select} shows how the value of each action is computed (and which action would be chosen by a greedy policy).

@@

\subsection{Motivation}\label{sec:motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
There are concrete reasons to use RL methods in this domain, beyond handling non-stationarity.
Misclassification is a serious problem, introducing \emph{collateral damage} in a DDoS prevention system.
RL's feedback-loop model lets us monitor flows \emph{after} an action is taken to perhaps allow forgiveness of mistakenly punished flows.
This relies upon a flow-by-flow view of the state space, but knowledge of the current state and last applied action detect whether punished flows fall off identically to legitimate flows.

%Which features might be best suited to this problem?
%?? Relevant features: aggregate network state (load at various points [this has been done, of course]), flow-specific measurements (upload/download ratio when bandwidth above threshold \cite{DBLP:conf/ndss/Rossow14}, packet inter-arrival times, etc.)
Other studies suggest that there are particularly useful features which make the task of online DDoS flow identification feasible.
Link loads observed at various locations suggest the overall health of a network \cite{DBLP:journals/eaai/MalialisK15}, and the ratio of correspondence between pair flows can suggest illegitimacy \cite{DBLP:conf/ndss/Rossow14}.
Volume-based statistics (counts, counts per duration, average packet sizes) have seen use in $k$-nearest neighbours classifiers trained to detect DDoS attacks \cite{DBLP:conf/dsn/LeeKSPY17}.
Most importantly, there is evidence showing that bandwidth expansion may cause behavioural changes which differ between benign and attack traffic \cite{DBLP:conf/ndss/KangGS16}.

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
We hypothesise that per-flow agent designs will offer better accuracy than current RL-based agents for DDoS mitigation.
We describe and justify our approach and algorithmic improvements, presenting two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

\subsection{System Design and Assumptions}
Agents are deployed in a network with a set of \emph{ingress/egress points} from its domain of control through which traffic can flow, and a set of protected \emph{destination nodes}.
These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, $U_s$.

@@

\subsubsection{Per-tile updates}
To adjust the value of each activated tile with a different magnitude and direction, we compute a different temporal difference value \emph{for each tiling}.
This corrects tiles which contribute too much or too little value without disturbing estimates which are almost correct.
While we make use of all tiles' action value estimates when making decisions, each tiling is updated using only its own contribution, allowing us to increase $\alpha$ without divergence.
Compared to the standard algorithm, effective policies are reached sooner.

@@

\begin{figure}
	\centering
	\resizebox{0.65\linewidth}{!}{
		\begin{tikzpicture}[
		texts/.style = {text=black},
		labeltexts/.style = {text=gray},
		treeline/.style = {draw=uofgburgundy},
		treenode/.style = {texts, circle, centered, fill=white, treeline},
		load/.style = {fill=uofgcobalt},
		loadline/.style = {draw=uofgcobalt, line width=0.75mm},
		loadhide/.style = {fill=uofgcobalt!40!white},
		external/.style = {fill=uofgrust},
		externalhide/.style = {fill=uofgrust!40!white},
		hideline/.style = {draw=uofgsandstone!40!white},
		hidenode/.style = {treenode, hideline},
		grow'=right
		]
		\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
		\node [treenode, loadhide, right = of mainagent] (inner0) {};
		\node [treenode, loadhide, above right = 0.2cm and 1cm of inner0] (inner1) {};
		\node [hidenode, below right = 0.2cm and 1cm of inner0] (inner2) {};
		\node [treenode, loadhide, below right = 0.2cm and 1cm of inner1] (inner3) {};
		\node [treenode, loadhide, right = of inner3] (inner4) {};
		\node [treenode, loadhide, above right = 0.2cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
		\node [hidenode, below right = 0.2cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};

		\node [hidenode, above left = 0.2cm and 0.5cm of inner1, label={[labeltexts]above:Agent 2}] (agent2) {};
		\node [hidenode, below left = 0.2cm and 0.5cm of inner2, label={[labeltexts]below:Agent 3}] (agent3) {};

@@

Our state space combines global state (network load) with per-flow measurements.
Each is tile-coded with 8 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
Features are tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.

%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}): the bandwidth measurements regularly received from monitors in the environment.
For any flow, an agent computes the path it would take through the network.
The incoming load along the first hop, last hop, and tertiles of the path are then tile-coded together.
In the event that this path is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
\Cref{fig:global-state-path} illustrates the process.
Global state built in this way is compatible with multipath, multi-destination networks, offering support for diverse deployment environments from endpoint servers up to transit ASes.
Computing a flow's path is not computationally expensive: multipath routing is fast since typical ECMP routing algorithms simply hash a packet's 5-tuple and are deterministic to provide consistent quality-of-service.

@@

\Cref{fig:udp-feature-plots} shows the effectiveness of per-flow features for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), on a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
The plots demonstrate that different protocols and traffic classes are best defended by different features---every feature presented has value in a complete model.
Packet-level and per-window statistics are at their based when combined with the last action for both UDP-like and TCP-like traffic (omitted).
All features converge to their highest-observed performance within around \num{4000} timesteps.
In general, some of the most effective features are global state, mean IAT, mean inbound packet size and $\Delta$ rates.
The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have no correlation with the class of traffic a flow contains.

@@

While our use and definition of $g(\cdot)$ appears nebulous, there are many ways to infer this quantity in practice.
End-host servers may make use of canary flows or active measurement schemes, and can employ quality-of-experience metrics in the case of VoIP services such as lost packets, reorderings, and jitter.
ASes and transit networks may make use of reports received from downstream, i.e. over the \emph{DDoS Open Threat Signalling} (DOTS) protocol \cite{ietf-dots-use-cases-17}.
We leave the use of such data as future work.
Even without online rewards, a well-trained agent needs only to follow the greedy policy it has learned from offline (simulated) training.

@@

\subsection{Action Space}
When monitoring a flow, an agent uses its state vector to decide which portion of its \emph{inbound} traffic should be dropped.
%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
We choose to drop packets ()rather than impose traffic limits) to have a discrete action space without prior knowledge of traffic characteristics.
Furthermore, we need not consider burstiness, fairness and parameters such as bucket sizes which could require tuning.
We offer two models, which differ only on how to choose $p$.
The intuition behind these designs is that an agent may directly choose the level of packet drop at the expense of a larger action space (\emph{Instant control}), or that we may reduce the action set (and learn a policy within fewer actions) by making use of domain knowledge (\emph{Guarded control}).

\subsubsection{Instant control}
Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$.
These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).

\subsubsection{Guarded control}
\Textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations, constraining how an agent treats each flow via simple finite state automata: restricting $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.

We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limits (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.
As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found the setting $\gamma = 0.8$ to be the most effective choice during exploratory testing.

To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$), chosen due to TCP's higher prevalence.
An agent must now choose to punish a flow multiple times in succession to remove it, reducing variance while allowing an agent to see how a host reacts to controlled conditions.
While beneficial for legitimate flows who behave similarly to attack traffic, this introduces a lag of an extra two actions in blocking illegitimate flows.

\subsubsection{Risks}
Our mode of action means that each agent carries a risk of introducing collateral damage into the network.
This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p\ne0$ in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
This weakness remains in modern flavours of TCP---TCP Cubic has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets and \SIrange{77}{84}{\percent} of data volume\footnote{Repository link removed for double-blind reviewing.}.
%Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
Furthermore, applying a single rule to many flows (as in \textcite{DBLP:journals/eaai/MalialisK15}) increases the likelihood of collateral damage.
When considering that many hosts have an especially adverse reaction to our main means of control, flow-level granularity becomes an obvious choice.
